\ttitle

### 2013-10-24 Appendices/AppendixA-summaries.tex
WIP
Added the TODO command

The dual problem maximizes this function under positivity constraints and one equality constraint:
*** TODO: fix equation (now commented) ***
\begin{equation}
\begin{aligned}
  % \alpha_1^*, \dots, \alpha_n^* = \argmax\limits_{\alpha_1, \dots, \alpha_n} - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i \\
  \text{subject to } \alpha_i \geq 0, 1 \leq i \leq n \text{ and } \sum\limits_{i=1}^n \alpha_i y_i = 0
\end{aligned}
\end{equation}

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci \cite{camci2010change}
About segmentation with SVMs.
Will be main material for section \ref{svm} about SVMs.

>>>>

The dual problem maximizes this function under positivity constraints and one equality constraint:
\TODO{fix equation (now commented)}
\begin{equation}
\begin{aligned}
  % \alpha_1^*, \dots, \alpha_n^* = \argmax\limits_{\alpha_1, \dots, \alpha_n} - \frac{1}{2} \sum\limits_{i=i}^n \sum\limits_{j=1}^n \alpha_i \alpha_j y_i y_j \vectorsym{x}_i \cdot \vectorsym{x}_j + \sum\limits_{i=1}^n \alpha_i \\
  \text{subject to } \alpha_i \geq 0, 1 \leq i \leq n \text{ and } \sum\limits_{i=1}^n \alpha_i y_i = 0
\end{aligned}
\end{equation}

@@

\subsection{Change Point Detection In Time Series Data Using Support Vectors}
Paper by Fatih Camci \cite{camci2010change}
About segmentation with SVMs.
Will be main material for section \ref{sec:svm_in_change_detection} about SVMs.

### 2013-10-24 Appendices/AppendixB-camci.tex
Added the TODO command

*** \emph{Thus methods of data descriptions should be compared (which are more general) and not only density estimation?} ***

@@

*** REVIEW zoeken: change points in time series ***

@@

*** TODO: make a clear distinction between model types; similarity (distance based) or model based, as in \cite{yin2008sensor} ***

@@

*** TODO: read \cite{mavroforakis2006geometric} and \cite{bennett2000duality} on geometry of SVMs ***

@@

*** Note: in \cite{mavroforakis2006geometric} it is explained why the inner-product between two vectors is a logical choice for the distance/similarity measure. ***

@@

\subsection{Novelty/outlier detection}
*** Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies. ***
\begin{itemize}
  \item One-class classification is referred to as ``Type 3'', with \emph{semi-supervised recognition or detection}.
  \item It explains why one-class can be beneficial over type-2 where negative examples needs to be provided.
  \item We are interested in type-3, so compare SVM with the others stated in this survey regarding type-3.
  \item Often refers to the convex hull of the data set. Link with geometric approach as in \cite{bennett2000duality,mavroforakis2006geometric}?
  \item It states that PCA and regression techniques are linear models and thus often are too simple for practical applications. SVMs try to find a hyperplane in higher dimensional space; linear models to implement complex class boundaries. It refers to \cite{tax1999support}.
\end{itemize}

>>>>

\TODO{\emph{Thus methods of data descriptions should be compared (which are more general) and not only density estimation?}}

@@

\TODO{REVIEW zoeken: change points in time series}

@@

\TODO{make a clear distinction between model types; similarity (distance based) or model based, as in \cite{yin2008sensor}}

@@

\TODO{read \cite{mavroforakis2006geometric} and \cite{bennett2000duality} on geometry of SVMs}

@@

\TODO{Note: in \cite{mavroforakis2006geometric} it is explained why the inner-product between two vectors is a logical choice for the distance/similarity measure.}

@@

\subsection{Novelty/outlier detection}
\TODO{Read \cite{hodge2004survey} ``A survey of outlier detection methodologies'' to compare with other methodologies.}
\begin{itemize}
  \item One-class classification is referred to as ``Type 3'', with \emph{semi-supervised recognition or detection}.
  \item It explains why one-class can be beneficial over type-2 where negative examples needs to be provided.
  \item We are interested in type-3, so compare SVM with the others stated in this survey regarding type-3.
  \item Often refers to the convex hull of the data set. Link with geometric approach as in \cite{bennett2000duality,mavroforakis2006geometric}?
  \item It states that PCA and regression techniques are linear models and thus often are too simple for practical applications. SVMs try to find a hyperplane in higher dimensional space; linear models to implement complex class boundaries. It refers to \cite{tax1999support}.
\end{itemize}

### 2013-10-24 Chapters/Chapter2.tex
Overview of papers for literature review 'Segmentation'
WIP
Added the TODO command

\section{Statistical framework}\label{statistical-framework}
Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... *** list and refs ***.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

@@

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
Section \ref{svm} discusses the \gls{svm}-method.
*** Put emphasis that this is the source of inspiration for the chosen method? ***

@@

*** Maybe add dimensionality reduction, but for now leave out ***

@@

>>>>

\section{Statistical framework}\label{statistical-framework}
Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

@@

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} takes an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
Section \ref{sec:svm_in_change_detection} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

@@

\TODO{Maybe add dimensionality reduction, but for now leave out}

@@

%----------------------------------------------------------------------------------------
% Temporal Segmentation
\input{./Chapters/Chapter2/temporal_segmentation.tex}

### 2013-10-24 Chapters/Chapter2/temporal_segmentation.tex
WIP
Added the TODO command

\subsection{Segmentation}\label{subsec:segmentation}
*** This subsection is mainly from previous draft version ***

*** TODO: Create compact notation, one sentence per paper max ***

@@

Many methods obtain an implicit segmentation as a result of classification over a sliding window *** add refs ***.
The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.

@@

\subsection{Change detection}\label{subsec:change_detection}
*** change order of this and temporal segmentation sections? So first change, then segmentation? *** \\
Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
Although these techniques are closely related, there is a subtle difference.
In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.

The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian *** Bayesian What? ***.
Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
Using an iterative approach, all possible change points are considered.
The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

*** TODO: move CUSUM based techniques to this subsection ***

>>>>

\subsection{Segmentation}\label{subsec:segmentation}
\TODO{This subsection is mainly from previous draft version}

\TODO{Create compact notation, one sentence per paper max}

@@

Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.

@@

\subsection{Change detection}\label{subsec:change_detection}
\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
Although these techniques are closely related, there is a subtle difference.
In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.

The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
Using an iterative approach, all possible change points are considered.
The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

Section \ref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

\TODO{move CUSUM based techniques to this subsection}

### 2013-10-24 Chapters/Chapter3.tex
Added the TODO command

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter \ref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In Section \ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
*** TODO: add \ref{sec:svm_parameters} ***
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in Section \ref{sec:quality_metrics}.

>>>>

This chapter discusses the concepts and algorithms that will be used as a basis for the proposed method, as discussed in Chapter \ref{Chapter4}.
The first sections formulates the problem of change detection and relates it to outlier and novelty detection.
It transforms the traditional problem to change detection for times series data.
In Section \ref{sec:one_class_classification} the problem of \acrlong{occ} is discussed and shows a number of implementations.
Two \acrlong{svm}-based methods, \gls{nu-svm} and \gls{svdd}, are further disussed in section \ref{sec:one_class_svm}.
\TODO{add \ref{sec:svm_parameters}}
For the application to the problem of \acrlong{har}, the characteristics of the collected sensor data is discussed in \ref{sec:sensor_data_characteristics}.
Finally a theoretic set of quality metrics is discussed in Section \ref{sec:quality_metrics}.

### 2013-10-24 Chapters/Chapter3/change_detection_time_series.tex
Added the TODO command

\subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
*** Give the problem of change definition. Follow definition of Camci \cite{camci2010change} ***

*** Note: the following block is copied from \ref{sec:method_change_indication} *** \\
The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
The algorithm searches for \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

@@

*** TODO: add more formal definition or analysis of change detection ***

@@

% \subsection{M-estimators}
% *** Move this section to somewhere in the end; is specification/extension, not base material ***
% To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
% The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
% Methods have been proposed to apply robust statistics in the form of M-estimators to \glspl{svm}, such as \cite{choi2009least,chen2004m,suykens1999least}.

% *** Question: do slack-variables (allowance of outliers) make SVM robust? ***

>>>>

\subsection{Problem formulation}\label{subsec:change_detection_problem_formulation}
\TODO{Give the problem of change definition. Follow definition of Camci \cite{camci2010change}}

\TODO{Note: the following block is copied from \ref{sec:method_change_indication}} \\
The problem of change point detection can be formulated using different type of models, as discussed in \ref{subsec:segmentation}.
The methods by Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} use the following formulation for change detection, which we will also use for our problem formulation.
The algorithm searches for \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kl} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

@@

\TODO{add more formal definition or analysis of change detection}

@@

% \subsection{M-estimators}
% \TODO{Move this section to somewhere in the end; is specification/extension, not base material}
% To make a method less sensitive to outliers (in the training data) techniques from robust statistics are applied.
% The term \emph{robustness} has many interpretations, one of them that it ``signifies insensi­tivity to small deviations from the [prior] assumptions [about the underlying situation]'', according to Huber \cite{huber2009robust}.
% Methods have been proposed to apply robust statistics in the form of M-estimators to \glspl{svm}, such as \cite{choi2009least,chen2004m,suykens1999least}.

% \TODO{Question: do slack-variables (allowance of outliers) make SVM robust?}

### 2013-10-24 Chapters/Chapter3/one_class_classification.tex
Added the TODO command

*** TODO: bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection ***

\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
*** Introduce problem of \gls{occ}. Relate to two-class classification. ***

@@

The \gls{occ} algorithms have been applied to a wide range of applications.
The first is, obviously, outlier detection of objects which do not resemble the bulk of the training data.
It can be a goal by itself and can be used as a filtering mechanism for other data processing methods.
Often methods that rely on data-characteristics are sensitive for remote regions in the data set.
Using \gls{occ} these remote regions can be removed from the data set.
A second application is for the problem as described above, in which only data from a single target class is available.
When this problems originates from \eg a monitoring process, the \gls{occ} is able to recognize abnormal system states, without the need to create (or simulate) faulty states beforehand.
The final possible application given by Tax \cite{tax2001one} is the comparison of two data sets.
By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
This will result in a similarity-measure.
It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
*** Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence? ***

@@

In this Section we have discussed the problem of \gls{occ} and different kind of implementations.
An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches.
*** TODO: add refs for good results of OC-SVM***
In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax2004support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.

>>>>

\TODO{bridge that explains why \gls{occ} is interesting, relating it to outlier/change detection}

\subsection{Problem formulation}\label{subsec:occ-problem-formulation}
\TODO{Introduce problem of \gls{occ}. Relate to two-class classification.}

@@

The \gls{occ} algorithms have been applied to a wide range of applications.
The first is, obviously, outlier detection of objects which do not resemble the bulk of the training data.
It can be a goal by itself and can be used as a filtering mechanism for other data processing methods.
Often methods that rely on data-characteristics are sensitive for remote regions in the data set.
Using \gls{occ} these remote regions can be removed from the data set.
A second application is for the problem as described above, in which only data from a single target class is available.
When this problems originates from \eg a monitoring process, the \gls{occ} is able to recognize abnormal system states, without the need to create (or simulate) faulty states beforehand.
The final possible application given by Tax \cite{tax2001one} is the comparison of two data sets.
By constructing a \gls{occ}-classifier using a training data set, one can compare that set to a new data set.
This will result in a similarity-measure.
It is easy to see here a relation with similiary measures over \glspl{pdf} and the \gls{kl} divergence. \\
\TODO{Add reference to \ref{subsec:change_detection_problem_formulation} which also mentions \gls{kl} divergence?}

@@

In this Section we have discussed the problem of \gls{occ} and different kind of implementations.
An often used implementation is the \gls{svm}-based method \cite{noumir2012simple}, since is shows good performance in comparitive researches.
\TODO{Add refs for good results of OC-SVM}
In the following section (\ref{sec:one_class_svm}) two implementations of \gls{oc-svm} will be discussed, the \gls{svdd} method of Tax and Duin \cite{tax2004support} and the \gls{nu-svm}-algoritm by Sch\"olkopf.

### 2013-10-24 Chapters/Chapter3/one_class_svm.tex
Added the TODO command

*** TODO: better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}. ***

This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
\\ *** TODO: Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
More in Section 9.3 of \cite{cherkassky2007learning}. *** \\
To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

*** TODO: Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}. ***

@@

The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions *** ref needed ***:
\begin{equation}
  K(x, x') = \operatorname{exp} \left( - \frac{ \lVert x - x' \rVert ^2}{2 \sigma^2 } \right),
\end{equation}
where $\sigma \in R$ is a kernel parameter and $\lVert x - x' \rVert$ is the dissimilarity measure expressed in Euclidean distance.

@@

*** Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning} ***

@@

*** Maybe add section about ``Change indication'' rather than ``detection''? ***

@@

*** These papers need to be referenced and included ***

>>>>

\TODO{better format of above formulas. Check with notation and formulas of chapter 9 of \cite{cherkassky2007learning}.}

This minimization problem can be solved (using \gls{qp}) and transformed to its Lagrange dual formulation.
\\ \TODO{Add explanation why to use dual, which is: because the problem formulation scales with the number of samples $n$ and not the dimesionality $d$ of the samples.
More in Section 9.3 of \cite{cherkassky2007learning}.} \\
To do so, the Lagrange multipliers $a_i >= 0$ are introduced and the decision function becomes:
\begin{equation}\label{eq:svm_lagrange}
  f(x) = \operatorname{sgn}( \sum_{i=1}^n \alpha_i y_i K(x, x_i) + b),
\end{equation}
where $K(x, x_i) = \phi(x)^T\phi(x_i)$ (which is further discussed in section \ref{subsec:kernel_trick}).
Here every data point $\mathcal{I}$ for which $a_i > 0$ is weighted in the decision function and thus ``supports'' the classification machine: hence the name ``\acrlong{svm}''.
Since it is shown that under certain circumstances \glspl{svm} show an equality to sparse representations \cite{girosi1998equivalence,smola1998connection}, there will often be relatively few Lagrange multipliers with a non-zero value.

\TODO{Add more on sparsity and good results, because of sparse representations, even in case of data of high dimensionality. e.g. \cite{cherkassky2007learning}.}

@@

The kernel function $K$ can have different forms, such as linear, polynomial and sigmoidal but the most used (and flexible) form is the Gaussian \gls{rbf}.
As Smola \etal \cite{smola1998connection} state, this Gaussian kernel yields good performance, especially when no assumptions can be made about the data.
The kernel maps input space $\mathcal{I}$ to the feature space $\mathcal{F}$ which is a Hilbert Space of infinite dimensions \TODO{ref needed}:
\begin{equation}
  K(x, x') = \operatorname{exp} \left( - \frac{ \lVert x - x' \rVert ^2}{2 \sigma^2 } \right),
\end{equation}
where $\sigma \in R$ is a kernel parameter and $\lVert x - x' \rVert$ is the dissimilarity measure expressed in Euclidean distance.

@@

\TODO{Add something about robustness, discusses in section 9.2 from \cite{cherkassky2007learning}}

@@

\TODO{Maybe add section about ``Change indication'' rather than ``detection''?}

@@

\TODO{These papers need to be referenced and included}

### 2013-10-24 Chapters/Chapter4/change_indication.tex
Added the TODO command

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}
*** TODO: change section title ***
*** TODO: ***
\emph{Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
Looks like my proposed methods, in that is combines outliers and gives a score to change points.}

@@

The problem statement and formal definition, following Tackeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
The algorithm needs to find \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
This is in line with the search of changes in activities, since we are only interested in different activities (which are represented by sudden changes) instead of changes within an activity.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kliep} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.

@@

*** TODO: What is my contribution compared to Camci? Distance of outliers to hypersphere? ***

>>>>

% !TEX root = ../../main.tex
\section{Change Indication}\label{sec:method_change_indication}
\TODO{change section title}
\TODO{***}
\emph{Follow methodology of ``A unifying framework for detecting outliers and change points from time series'' \cite{takeuchi2006unifying}.
It creates a two-stage process of first searching for outliers, and then using the ``outlier-score'' to find (sudden) change points, by a weighted average of a moving window.
That eventual score can be thresholded (as in the paper) or processed with something like CUSUM (proposal).
Looks like my proposed methods, in that is combines outliers and gives a score to change points.}

@@

\TODO{the following paragraph is copied into \ref{subsec:change_detection_problem_formulation}. Remove here? Or shorten?}\\
The problem statement and formal definition, following Takeuchi and Yamanishi \cite{takeuchi2006unifying} and Camci \cite{camci2010change} is the following.
The algorithm needs to find \emph{sudden} changes in the time series data.
In other words, slowly changing properties in the data are not considered to be changes.
This is in line with the search of changes in activities, since we are only interested in different activities (which are represented by sudden changes) instead of changes within an activity.
Considered a time series $x_1 x_1 \dots$, which is drawn from a stochastic process $p$.
Each $x_t$ (t = 1, 2, \dots) is a $d$-dimensional real valued vector and $p$ a probability density function of the sequence $x_1 x_2 \dots$.
Assume $p$ can be decomposed in two different \gls{iid} stationary stochastic processes $p^1$ and $p^2$ and are one-dimensional Gaussian density functions.
For a time point $a$ data points for which $t < a$ are drawn from $p^1 = N(\mu_1, \sigma_1^2)$ and for $t \geq a$ from $p^2 = N(\mu_2, \sigma_a^2)$.
If $p^1$ and $p^2$ are different, then the time point $t = a$ is a \emph{change point}.
In \cite{takeuchi2006unifying} the similarity between the stochastic processes are expressed by the \gls{kliep} divergence $D(p^2||p^1)$.
The problem with this measure is that, as the authors conclude and Camci discusses, it is not able to detect a change by decrease in variance.
\\\TODO{End copied paragraph}

@@

\TODO{What is my contribution compared to Camci? Distance of outliers to hypersphere?}

### 2013-10-24 Chapters/old_classification/Chapter1/lit_rev_segm_and_class.tex
Added the TODO command

During the last ten years much research is performed, as a result of smaller and cheaper measuring devices. *** cite needed ***.
With the introduction of smartphones with sensory hardware the research has shifted to these devices, as they create a more natural and ubiquitous setting for the test subject. *** cite needed ***.

@@

Very much used in machine learning problems are neural networks. *** cite needed ***
Although these kind of networks typically perform well on spatial-typed problems *** cite needed ***, adjustments have been made to apply them for temporal classification.
The methods of \cite{yang2008using, kwapisz2011activity} create a multi-level neural network to generate complex discriminators as activity classifiers.
A variant of neural networks known as Kohonen Self Organizing Maps is used in \cite{krause2003unsupervised} to create a vector codebook, which is further processed by a $k$-means clustering.

@@

Some researchers have made their used data set explicitly public available and other data sets are just available, providing valuable labeled activity patterns. *** This should be in experiments setup? ***
*** List here the datasets used ***

>>>>

During the last ten years much research is performed, as a result of smaller and cheaper measuring devices. \TODO{cite needed}.
With the introduction of smartphones with sensory hardware the research has shifted to these devices, as they create a more natural and ubiquitous setting for the test subject. \TODO{cite needed}.

@@

Very much used in machine learning problems are neural networks. \TODO{cite needed}
Although these kind of networks typically perform well on spatial-typed problems \TODO{cite needed}, adjustments have been made to apply them for temporal classification.
The methods of \cite{yang2008using, kwapisz2011activity} create a multi-level neural network to generate complex discriminators as activity classifiers.
A variant of neural networks known as Kohonen Self Organizing Maps is used in \cite{krause2003unsupervised} to create a vector codebook, which is further processed by a $k$-means clustering.

@@

Some researchers have made their used data set explicitly public available and other data sets are just available, providing valuable labeled activity patterns. \TODO{This should be in experiments setup?}
\TODO{List here the datasets used}

### 2013-10-24 Chapters/old_classification/Chapter1/lit_rev_temp_segm.tex
Create outlines for new focus: temporal segmentation
Added the TODO command

*** Piecewise linear approximation ***

@@

The time series data from a sensor can be considered as a stochastic process *** is this a correct term? ***.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The CUSUM-method relies on the log-likelihood ratio to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}) directly, without explicit estimation of the densities.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

@@

>>>>

\TODO{Piecewise linear approximation}

@@

The time series data from a sensor can be considered as a stochastic process \TODO{is this a correct term?}.
Probabilistic models can be constructed on that signal, yielding in probabilistic and Bayesian based segmentation methods.
The CUSUM-method relies on the log-likelihood ratio to measure the difference between two distributions.
To calculate the ratio, the probability density functions need to be calculated.
The method of Kawahara \etal \cite{kawahara2009change} proposes to estimate the ratio of probability densities (known as the \emph{importance}) directly, without explicit estimation of the densities.
They claim this results in a robuster approach for real-world scenarios.
Although this is a model-based method, no strong assumptions (parameter settings) are made on the models.

@@

\TODO{TODO} aan het einde van de lopende tekst richting-zinnen maken, die het einde aanduiden of een erg beknopte samenvatting geven / conclusie van literatuuronderzoek.
